### 2023-01-22 chapter1/chapter.tex
.
post-reviews, pre supervisor feedback

A stochastic process is \emph{stationary} if, for any \(n\), \(t_0<t_1<...<t_n\) and any \(t\), then \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution. That is, 
\begin{align}\label{eqn: stationary}
	&\nonumber\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
	\\&= \mathbb P(X(t_n+t)\in E_n,X(t_{n-1}+t)\in E_{n-1},...,X(t_0+t)\in E_0),
\end{align}
for any \(t,t_0,...,t_n\in\mathcal T\) with \(t_0+t,...,t_n+t\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\). A random variable, \(\tau\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\tau\in\sigma(X(s), s\leq t),\) for all \(t\in T\), where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s)\}_{s\leq t}\).

As in \cite[Chapter~1.2]{MEinAP}, we call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
\begin{align}\label{eqn: Markov property}
	\nonumber &\mathbb P(X(n+1)=j \mid X(n)=i,X(n-1)=i_{n-1},...,X(0)=i_0) 
	\\&= \mathbb P(X(n+1)=j \mid X(n)=i) 
\end{align}
for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and referred to (\ref{eqn: Markov property}) as the \emph{Markov property}. The process \(\{X(n)\}_{n\in\{0,1,2,...\}}\) is said to be time-homogeneous if 
\(\mathbb P(X(n+1)=j \mid X(n)=i)\) does not depend on \(n\). The probabilities \(\mathbb P(X(n+1)=j \mid X(n)=i)=:p_{ij}\) are the \emph{transition probabilities}, and \(\bs P=[p_{ij}]_{i,j\in\calS}\) is the \emph{transition matrix}, of the Markov chain. The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(n)\}_{n\in\{0,1,2,...\}}\), on the event that \(\{\tau<\infty\}\), then 
\begin{align}\label{eqn: strong Markov property}
	\mathbb P(X(\tau+1)=j \mid \sigma(X(0),X(1),...,X(\tau))) = \mathbb P(X(\tau+1)=j \mid X(\tau)).
\end{align}

@@

An \emph{operator} is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space). That is, let \(V\) and \(W\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\to W\) is an operator. The set \(\mathcal D(\mathbb B)\subseteq V\) is the domain of the operator, and is not necessarily all of \(V\). When specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to W\).

Let \(c\in \mathbb R\) (\(\mathbb R\) is the reals), \(v\in V\) and \(w\in W\). Then \(\mathbb B\) is a \emph{linear} operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\). We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_W}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_W\) are the norms defined on \(V\) and \(W\), respectively.  An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty\).

@@

For a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}:=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\). It can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), which implies that \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup. As another example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega(\mathbb R) \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\). Then \(\mathbb T(t)f(x)=\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is the Taylor series of \(f\) about the point \(x\), and we see that the operator \(\mathbb T(t)f(x)=f(x-t)\) is the shift operator. The semigroup property follows since 
\[\mathbb T(t)\mathbb T(s)f(x)=\mathbb T(t)f(x-s)=f(x-s-t)=\mathbb T(t+s)f(x).\]

@@

\subsubsection{Markov processes and semigroups}
Semigroups arise naturally in Markov processes. Let \(\{X(t):t\geq 0\}\) be a Markov process with state space \(\mathcal S\). The Markov property states that, given \(X(t)\), then for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\). That is, \[P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=P(X(t+s)\in\mathcal{E} \mid X(t)).\] 
If \(P(X(t+s)\in\mathcal{E} \mid X(t))=P(X(s)\in\mathcal{E} \mid X(0))\), are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.

@@

Consider the expectation
\[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) P(X(t)\in\wrt y\mid X(0)=x),\]
where \(f\) is some bounded function, \(|f|<F\). We can think of expectation as an operator which is acting on the function \(f\). The expectation depends on \(t\) and \(x\) as these define the distribution with which we are taking the expectation, hence, let us write 
\[\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x].\]
The result of the operator acting on \(f\), \(\mathbb T(t)f\), is another function (a function of the initial point \(x\)).

By the Markov property, for any \(s\in[0,t]\), 
\begin{align*}
	\mathbb T(t)f(x)&=\int_{y\in\calS} f(y) P(X(t)\in\wrt y \mid X(0)=x)
	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}f(y) P(X(t)\in\wrt y\mid X(s)=z) P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\mathbb T(s) (\mathbb T(t-s)f)(x),
\end{align*}
where the third equality holds by time-homogeneity. Hence, \(\{\mathbb T(t)(\cdot)\}_{t\geq 0}\) is an operator semigroup. Moreover, since \(f\) is bounded, then \(\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x] \leq F\), which means that \(\{\mathbb T(t)\}\) is a contraction semigroup.

@@

% Under certain conditions, a semigroup, (and hence also a Feller process), can be characterised by its generator, \(B\) defined by;
% \[Bf:=\lim_{t\to 0^+} \frac{1}{t}(K(t)-I)f,\]
% on the set of functions \(f\) for which this limit exists. This set is known as the domain of the generator \(\mathcal D(B)\), which is not necessarily all of \(L\). The domain of \(B\) is a very important part of its definition, and some authors highlight this by writing the generator as the pair \((B,\mathcal D(B))\).

@@

\section{Laplace Transforms}
The Laplace transform is an important tool in many areas of mathematics, and is particularly useful in the analysis of fluid queues and fluid-fluid queues. Furthermore, in Chapters~\ref{sec: conv} and~\ref{ch: global results} we work with {Laplace transforms} to prove convergence of the QBD-RAP scheme.

For a measure \( \mu\), defined on the Borel sets of \([0,\infty)\), we define the Laplace transform of \(\mu\) to be
\[\widehat \mu(\lambda) = \mathcal L(\mu)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}\wrt \mu,\]
where the \emph{region of convergence} is the set of values of \(\lambda\in\mathbb R\) such that the integral is finite. When \(\mu\) has a density, \(v\), then the Laplace transform is 
\[\widehat \mu(\lambda) = \widehat v(\lambda) = \mathcal L(v)(\lambda) = \int_{t=0}^\infty e^{-\lambda t}v(x)\wrt x.\]
When \(\mu\) is the probability measure associated with a random variable, \(Z\), say, then we may write 
\[\widehat \mu(\lambda)=\mathbb E[e^{-\lambda Z}],\]
and the region of convergence is at least \([0,\infty)\). 
Further, letting \(E^\lambda\) be an exponentially distributed random variable with rate \(\lambda\) and noting that \(\mathbb P(E^\lambda >t ) = e^{-\lambda t}\), then 
\[\widehat \mu(\lambda)=\mathbb P(Z<E^\lambda),\]
which gives a probabilistic interpretation of the Laplace transform. That is, the Laplace transform with parameter \(\lambda >0\) is the probability that \(Z\) occurs before \(E^\lambda\), an independent random exponential time with rate \(\lambda\), occurs.

A convenient property of the Laplace transform which we utilise is the Convolution Theorem. 
\begin{thm}[Convolution Theorem]
	Let \(f,g: [0,\infty) \to \mathbb R\) be integrable functions, then 
	\[\mathcal L\left(\int_{u=0}^tf(u)g(t-u)\wrt u\right)(\lambda) = \mathcal L\left(f\right)\cdot \mathcal L\left(g\right).\]
\end{thm}
The Convolution Theorem states that the Laplace transform of the convolution, given by \(\displaystyle \int_{u=0}^tf(u)g(t-u)\wrt u\), is equal to the product of the Laplace transform of \(f\) and \(g\).

The Laplace transform is unique in the sense that, if \(\mu\) and \(\nu\) are two measures on the Borel sets of \([0,\infty)\) and 
\[\widehat \mu(\lambda) = \widehat \nu(\lambda)\] 
for all \(\lambda > a\) with \(a<\infty\), then \(\mu\) and \(\nu\) are the same. In terms of functions, \(f,g: [0,\infty) \to \mathbb R\), if 
\[\mathcal L(f)(\lambda) = \mathcal L(g)(\lambda),\]
for all \(\lambda > a\) with \(a<\infty\), and \(f\) and \(g\) are continuous, then \(f(t)=g(t)\) for all \(t\geq 0\). Without knowing \(f\) and \(g\) are continuous, then we can only claim that 
\(f(t)=g(t)\) for all \(t\geq 0, t\notin \mathcal N,\) where \(\mathcal N\) is a \emph{null set} with respect to Lebesgue measure.

\section{Fluid queues}\label{sec: fqs}
An unbounded fluid queue is a two-dimensional stochastic process which we denote by \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer. The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\), and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\). We assume that \(\bs T\) is {conservative} and time-homogeneous. Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\) which determine the rate at which \(\{X(t)\}\) moves.

@@

{Let's clarify some notation. We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some index set. Similarly, \(\bs u = (\bs u_h)_{h\in\mathcal H}\), is a row-vector defined by a collection of row-vectors \(\bs u_h\). The notation \(\bs u_m=(u_h)_{h\in\mathcal H_m}\) refers to the vector containing the subset of elements corresponding to \(\mathcal H_m\subseteq \mathcal H\). When the index set is empty, the resulting vector \(\bs u_m\) is a vector of dimension 0. In cases when there are two indices, we order the elements of the vector according to the first index, then the second; i.e.~\(\bs u = (u_{g}^h)_{g\in\mathcal G,h\in\mathcal H} = ((u_g^h)_{g\in\mathcal G})_{h\in\mathcal H}\). Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\). The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).}

@@

Upon hitting a boundary we suppose that, with probability \(p_{ij},\,i,j\in\mathcal S\), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(\mathrm{sign}(c_i)=\mathrm{sign}(c_j)\) or \(\mathrm{sign}(c_j)=0\) then the process is absorbed in the boundary, otherwise it immediately leaves the boundary. At a lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator 
\[\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_+\), at which time \(\{X(t)\}\) leaves the boundary. Similarly, at an upper boundary if \(j\in\calS_0\cup\calS_+\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\) and the phase process continues to evolve according to the sub-generator 
\[\left[\begin{array}{cc} \bs T_{++} & \bs T_{+0} \\ \bs T_{0+} & \bs T_{00}  \end{array}\right],\]
until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_-\) at which time \(\{X(t)\}\) leaves the boundary. It is without loss of generality that we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=b>0\), respectively.

@@

Consider an unbounded fluid queue \(\{(\ddot X(t), \varphi(t))\}\). Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(\lambda)\) with elements \(\left[\bs \Psi_X(\lambda)\right]_{ij}\), \(i\in\calS_+\), \(j\in\calS_-\), given by
\begin{align}
	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j))\mid \ddot X(0)=z, \varphi(0)=i\right].
\end{align}
\(\left[\bs \Psi_X(\lambda)\right]_{ij}\) is the Laplace-Stieltjes transform of the time taken for the fluid queue to first return to level \(z\) and do so in phase \(j\in\calS_-\), given it started at level \(z\) in phase \(i\in\calS_+\). Define the in-out fluid level by $\beta_X(t) := \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in to or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and also define $\eta_X(y) := \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\). Further, let \(\bs H(\lambda,y)\) be the matrix with elements \(h_{ij}(\lambda,y)\), \(i,j\in\calS_+\cup\calS_-\), given by
\begin{align}
	\mathbb E\left[e^{-\lambda \eta_X(y)}1(\eta_X(y)<\infty,\varphi(\eta_X(y))=j)\mid \ddot X(0)=0, \varphi(0)=i\right],
\end{align}
which is the Laplace-Stieltjes transform of the time taken for \(y\) amount of fluid to flow in or out of the fluid queue and to be in phase \(j\) at this time, given the initial level of the fluid queue was \(0\) and the initial phase was \(i\).

>>>>

A stochastic process is \emph{stationary} if, for any \(n\), \(t_0<t_1<...<t_n\) and any \(t\), then \(X(t_n),X(t_{n-1}),...,X(t_0)\) and \(X(t_n+t),X(t_{n-1}+t),...,X(t_0+t)\) have the same distribution. That is, 
\begin{align}\label{eqn: stationary}
	&\nonumber\mathbb P(X(t_n)\in E_n,X(t_{n-1})\in E_{n-1},...,X(t_0)\in E_0) 
	\\&= \mathbb P(X(t_n+t)\in E_n,X(t_{n-1}+t)\in E_{n-1},...,X(t_0+t)\in E_0),
\end{align}
for any \(t,t_0,...,t_n\in\mathcal T\) with \(t_0+t,...,t_n+t\in\mathcal T\) and any \(E_0,...,E_n\subseteq \mathcal S\).

As in \cite[Section~2.1]{MEinAP}, a random variable, \(\tau\) taking values in \(\mathcal T\), is a \emph{stopping time} for the stochastic process \(\{X(t)\}_{t\in\mathcal T}\) if \(\{\tau\leq t\}\in\sigma(X(s), s\leq t),\) for all \(t\in T\), where \(\sigma(X(s), s\leq t)\) denotes the \(\sigma\)-algebra generated by \(\{X(s)\}_{s\leq t}\).

As in \cite[Section~1.2]{MEinAP}, we call \(\{X(n)\}_{n\in\{0,1,2,\dots\}}\) a \emph{discrete-time Markov chain} if \(\calS\) is countable and
\begin{align}\label{eqn: Markov property}
	\nonumber &\mathbb P(X(n+1)=j \mid X(n)=i,X(n-1)=i_{n-1},...,X(0)=i_0) 
	\\&= \mathbb P(X(n+1)=j \mid X(n)=i) 
\end{align}
for all \(n\in\{0,1,2,...\}\) and \(i,i_0,...,i_{n-1},j\in\calS,\) and referred to (\ref{eqn: Markov property}) as the \emph{Markov property}. The process \(\{X(n)\}_{n\in\{0,1,2,...\}}\) is said to be time-homogeneous if 
\(\mathbb P(X(n+1)=j \mid X(n)=i)\) does not depend on \(n\). The probabilities \(\mathbb P(X(n+1)=j \mid X(n)=i)=:p_{ij}\) are the \emph{transition probabilities}, and \(\bs P=[p_{ij}]_{i,j\in\calS}\) is the \emph{transition matrix}, of the Markov chain. The \emph{strong Markov property} states that for each stopping time \(\tau\) of the Markov chain \(\{X(n)\}_{n\in\{0,1,2,...\}}\), on the event that \(\{\tau<\infty\}\), then 
\begin{align}\label{eqn: strong Markov property}
	\mathbb P(X(\tau+1)=j \mid \sigma(X(0),X(1),...,X(\tau))) = \mathbb P(X(\tau+1)=j \mid X(\tau)).
\end{align}

@@

An \emph{operator} is a mapping which takes elements of a vector space and sends them to elements of a vector space (not necessarily the same space). That is, let \(V\) and \(U\) be normed vector spaces, then \(\mathbb B:\mathcal D(\mathbb B)\to U\) is an operator. The set \(\mathcal D(\mathbb B)\subseteq V\) is the domain of the operator, and is not necessarily all of \(V\). When specification of the domain is unnecessary we sometimes write \(\mathbb B:V\to U\).

Let \(c\in \mathbb R\) (\(\mathbb R\) is the reals), \(v\in V\) and \(w\in U\). Then \(\mathbb B\) is a \emph{linear} operator if \(\mathbb B(cv) = c(\mathbb Bv)\) and \(\mathbb B(v+w)=\mathbb Bv+\mathbb Bw\). We define the norm of an operator as \(||\mathbb B||=\sup_{v\in V}\cfrac{||\mathbb Bv||_U}{||v||_V}\), where \(||\cdot||_V\) and \(||\cdot||_U\) are the norms defined on \(V\) and \(U\), respectively.  An operator is said to be bounded if \(||\mathbb B||<M\) for some \(M<\infty\).

@@

For a bounded linear operator \(\mathbb B:\mathcal D(\mathbb B)\to \mathcal D(\mathbb B)\), define the operator exponential as \(e^{\mathbb Bt}=\sum_{k=0}^\infty\cfrac{1}{k!}t^k\mathbb B^k\) where, for \(f\in \mathcal D(\mathbb B)\), \(\mathbb B^kf\) is defined recursively by \(\mathbb B^k f = \mathbb B^{k-1} (\mathbb B f)\). It can be shown that \(e^{\mathbb B(t+s)}=e^{\mathbb Bt}e^{\mathbb Bs}\), and from the definition \(e^{\mathbb B0}=I\), which implies that \(\{e^{\mathbb Bt}:t\geq 0\}\) is a semigroup. As another example, let \(\mathbb B=\cfrac{d}{dx}:C^\omega(\mathbb R) \to C^\omega(\mathbb R)\), where \(C^\omega(\mathbb R)\) is the class of analytic functions on the whole of \(\mathbb R\). Then \(\mathbb T(t)f(x)=\sum_{k=0}^\infty \cfrac{1}{k!}t^k\cfrac{\wrt^k}{\wrt x^k}f(x)=f(x-t)\), and the sum converges since \(f\in C^\omega(\mathbb R)\); the series representation is the Taylor series of \(f\) about the point \(x\), and we see that the operator \(\mathbb T(t)f(x)=f(x-t)\) is the shift operator. The semigroup property follows since 
\[\mathbb T(t)\mathbb T(s)f(x)=\mathbb T(t)f(x-s)=f(x-s-t)=\mathbb T(t+s)f(x).\]

@@

\subsubsection{Markov processes and semigroups}
Semigroups arise naturally in Markov processes. Let \(\{X(t):t\geq 0\}\) be a Markov process with state space \(\mathcal S\). The Markov property states that, given \(X(t)\), then for \(s,t\geq 0\), \(X(t+s)\) is independent of \(X(u),\, 0\leq u<t\). That is, \[\mathbb P(X(t+s)\in\mathcal{E} \mid X(t), X(u), 0\leq u<t)=\mathbb P(X(t+s)\in\mathcal{E} \mid X(t)).\] 
If \(\mathbb P(X(t+s)\in\mathcal{E} \mid X(t))=\mathbb P(X(s)\in\mathcal{E} \mid X(0))\), are invariant under translation by \(t\), then \(X(t)\) is said to be time-homogeneous.

@@

Consider the expectation
\[\mathbb E[f(X(t))\mid X(0)=x]=\int_{y\in\calS} f(y) \mathbb P(X(t)\in\wrt y\mid X(0)=x),\]
where \(f\) is some bounded function, \(|f|<F\). We can think of expectation as an operator which is acting on the function \(f\). The expectation depends on \(t\) and \(x\) as these define the distribution with which we are taking the expectation, hence, let us write 
\[\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x].\]
The result of the operator acting on \(f\), \(\mathbb T(t)f\), is another function (a function of the initial point \(x\)).

By the Markov property, for any \(s\in[0,t]\), 
\begin{align*}
	\mathbb T(t)f(x)&=\int_{y\in\calS} f(y) \mathbb P(X(t)\in\wrt y \mid X(0)=x)
	\\&=\int_{y\in\calS}\int_{z\in\mathcal S}f(y) \mathbb P(X(t)\in\wrt y\mid X(s)=z) \mathbb P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\int_{z\in\mathcal S}\mathbb E[f(X(t))\mid X(s)=z] \mathbb P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\int_{z\in\mathcal S} \mathbb E[f(X(t-s))\mid X(0)=z]\mathbb P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\int_{z\in\mathcal S} (\mathbb T(t-s)f(z))\mathbb P(X(s)\in\wrt z\mid X(0)=x)
	\\&=\mathbb T(s) (\mathbb T(t-s)f)(x),
\end{align*}
where the third equality holds by time-homogeneity. Hence, \(\{\mathbb T(t)(\cdot)\}_{t\geq 0}\) is an operator semigroup. Moreover, since \(f\) is bounded, then \(\mathbb T(t)f(x) = \mathbb E[f(X(t))\mid X(0)=x] \leq F\), which means that \(\{\mathbb T(t)\}\) is a contraction semigroup.

@@

% Under certain conditions, a semigroup, (and hence also a Feller process), can be characterised by its generator, \(B\) defined by;
% \[Bf=\lim_{t\to 0^+} \frac{1}{t}(K(t)-I)f,\]
% on the set of functions \(f\) for which this limit exists. This set is known as the domain of the generator \(\mathcal D(B)\), which is not necessarily all of \(L\). The domain of \(B\) is a very important part of its definition, and some authors highlight this by writing the generator as the pair \((B,\mathcal D(B))\).

@@

\section{Fluid queues}\label{sec: fqs}
An unbounded fluid queue is a two-dimensional stochastic process which we denote by \(\{\ddot{\bs X}(t)\} = \{(\ddot X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process of the fluid queue, and \(\{\ddot X(t)\}_{t\geq0}\) is known as the level process or buffer. The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we assume to be \(\mathcal S=\{1,2,\dots,N\}\), and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\)\footnote{Let's clarify some notation. We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some index set. Similarly, \(\bs u = (\bs u_h)_{h\in\mathcal H}\), is a row-vector defined by a collection of row-vectors \(\bs u_h\). The notation \(\bs u_m=(u_h)_{h\in\mathcal H_m}\) refers to the vector containing the subset of elements corresponding to \(\mathcal H_m\subseteq \mathcal H\). When the index set is empty, the resulting vector \(\bs u_m\) is a vector of dimension 0. In cases when there are two indices, we order the elements of the vector according to the first index, then the second; i.e.~\(\bs u = (u_{g}^h)_{g\in\mathcal G,h\in\mathcal H} = ((u_g^h)_{g\in\mathcal G})_{h\in\mathcal H}\). Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\). The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) (square brackets) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\).}. We assume that \(\bs T\) is {conservative} and time-homogeneous. Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\) which determine the rate at which \(\{X(t)\}\) moves.

@@

@@

Upon hitting the lower (upper) boundary we suppose that, with probability \(\widecheck p_{ij},\,i,j\in\mathcal S\) (\(\widehat p_{ij},\,i,j\in\mathcal S\)), the phase process instantaneously transitions from phase \(i\) to phase \(j\) (note that we might have \(i=j\) i.e.~no transition) and if \(\mathrm{sign}(c_i)=\mathrm{sign}(c_j)\) or \(\mathrm{sign}(c_j)=0\) then the process is absorbed in the boundary, otherwise it immediately leaves the boundary. At the lower boundary, if \(j\in\calS_0\cup\calS_-\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\), and the phase process continues to evolve according to the sub-generator 
\[\left[\begin{array}{cc} \bs T_{--} & \bs T_{-0} \\ \bs T_{0-} & \bs T_{00}  \end{array}\right],\]
until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_+\), at which time \(\{X(t)\}\) leaves the boundary. Similarly, at the upper boundary if \(j\in\calS_0\cup\calS_+\), then \(\cfrac{\wrt}{\wrt t} X(t) = 0\) and the phase process continues to evolve according to the sub-generator 
\[\left[\begin{array}{cc} \bs T_{++} & \bs T_{+0} \\ \bs T_{0+} & \bs T_{00}  \end{array}\right],\]
until such a time that \(\{\varphi(t)\}\) transitions to a phase \(k\in\calS_-\) at which time \(\{X(t)\}\) leaves the boundary. It is without loss of generality that we assume the lower and upper boundaries (when present) are at \(x=0\) and \(x=b>0\), respectively.

@@

Consider an unbounded fluid queue \(\{(\ddot X(t), \varphi(t))\}\). Let \(\zeta_X(E)\) be the random variable which is the first hitting time of \(\{\ddot X(t)\}\) on the set \(E\) and define the matrix \(\bs \Psi_X(\lambda)\) with elements \(\left[\bs \Psi_X(\lambda)\right]_{ij}\), \(i\in\calS_+\), \(j\in\calS_-\), given by
\begin{align}
	\left[\bs \Psi_X(\lambda)\right]_{ij} = \mathbb E\left[ e^{-\zeta_X(z)\lambda} 1(\zeta_X(z)<\infty, \varphi(\zeta_X(z)=j))\mid \ddot X(0)=z, \varphi(0)=i\right].
\end{align}
\(\left[\bs \Psi_X(\lambda)\right]_{ij}\) is the Laplace-Stieltjes transform of the time taken for the fluid queue to first return to level \(z\) and do so in phase \(j\in\calS_-\), given it started at level \(z\) in phase \(i\in\calS_+\). Define the in-out fluid level by $\beta_X(t) = \int_0^t \left| c_{\varphi(z)} \right|  \wrt z$, which is the total amount of fluid to flow in to or out of the buffer \(\{\ddot X(t)\}\) by time \(t\), and also define $\eta_X(y) = \inf \{t > 0: \beta_X(t) = y\}$ as the first hitting time of the in-out process \(\{\beta_X(t)\}\) on level \(y\geq 0\). Further, let \(\bs H(\lambda,y)\) be the matrix with elements \(h_{ij}(\lambda,y)\), \(i,j\in\calS_+\cup\calS_-\), given by
\begin{align}
	\mathbb E\left[e^{-\lambda \eta_X(y)}1(\eta_X(y)<\infty,\varphi(\eta_X(y))=j)\mid \ddot X(0)=0, \varphi(0)=i\right],
\end{align}
which is the Laplace-Stieltjes transform of the time taken for \(y\) amount of fluid to flow in or out of the fluid queue and to be in phase \(j\) at this time, given the initial level of the fluid queue was \(0\) and the initial phase was \(i\).
